The polynomial representation of the type $A_{n - 1}$ rational Cherednik algebra in characteristic $p \mid n$

### 2016-03-28 submission3r2.tex
Rewrite ci proof

\newcommand{\xx}[1]{\textcolor{blue}{#1}}
\newcommand{\zb}{\overline{z}}
\newcommand{\cc}[1]{\overline{#1}}
\newcommand{\m}[1]{\left|#1\right|}
\newcommand{\Res}{\operatorname{Res}}
\newcommand{\p}[1]{\left(#1\right)}
\newcommand{\rank}{\operatorname{rank}}
\newcommand{\sgn}{\operatorname{sgn}}
\newcommand{\D}{\triangle}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\f}[1]{\sum_{n=-\infty}^\infty #1_ne^{inx}}
\newcommand{\fn}[1]{\sum_{n=-\infty}^\infty #1e^{inx}}
\newcommand{\BB}{\mathcal{B}}
\newcommand{\M}{\mathcal{M}}
\newcommand{\mm}[1]{\m{\m{#1}}}
\newcommand{\ang}[1]{\left\langle #1 \right\rangle}
\newcommand{\h}{\mathfrak{h}}
\newcommand{\GL}{{\rm GL}}
\newcommand{\HH}{\mathcal{H}}
\newcommand{\arxiv}[1]{\href{http://arxiv.org/abs/#1}{{\tt arXiv:#1}}}
\newcommand{\Sym}{\operatorname{Sym}}
\renewcommand{\char}{\text{char}}
\newcommand{\sspan}{\text{span}}
\newcommand{\wF}{\widetilde{F}}
\newcommand{\supp}{\text{supp}}

@@

The authors thank P. Etingof for suggesting the problem and for helpful discussions.  Some exploratory computations were done using Sage.  S.~D. was supported by the MIT Undergraduate Research Opportunities Program (UROP). Y.~S. was supported by a NSF Graduate Research Fellowship (NSF Grant \#1122374).  Both authors were also supported by NSF Grant DMS-1000113.

@@

\subsection{Proving $f_i$ are singular vectors}

We now show that $f_1, \ldots, f_n$ are singular vectors summing to $0$.

\begin{lemma} \label{lem:sum-zero}
For $c \neq 0$, we have $\sum_{i = 1}^n f_i = 0$.
\end{lemma}
\begin{proof}
By Lemma \ref{lem:dFdz}, we have that
\[
zF'(z) = z V(z) - \sum_{j = 1}^n \frac{cx_jz}{1 - x_j z} F(z) = z V(z) - c \sum_{j = 1}^n \frac{1}{1 - x_j z} F(z),
\]
where we subtract $ncF(z)$ in the second equality.  Notice that $[z^p]$ coefficient of the left side is $0$ because $[z^{p - 1}] F'(z) = 0$ and that $[z^p] zV(z) = 0$.  Therefore, we conclude that $c \sum_{i = 1}^n f_i = 0$, giving the desired.
\end{proof}

@@

Consider the ideal $I_c = \langle f_1, \ldots, f_{n - 1} \rangle \subset A$ generated by the $f_i$.  In this section, we will show that $A/I_c$ is a complete intersection.  Recall that for an ideal $I \subset k[X_1,\dots,X_m]$, the quotient ring $k[X_1,\dots,X_m]/I$ is a complete intersection if $I$ has a minimal set of generators $g_1, \ldots, g_l$ with $\dim k[X_1,\dots,X_m]/I = m - l$.

\begin{proposition}\label{prop:ci}
For generic $c$, the quotient $A/I_c$ is a complete intersection. 
\end{proposition}
\begin{proof}
By Proposition~\ref{prop:linind}, $I_c$ has a set of $n-1$ linearly independent and therefore minimal generators $f_1, \ldots, f_{n-1}$ in degree $p$.  We will now show that for generic $c$, if $\alpha_1,\dots,\alpha_{n} \in k$ satisfy $f_i(\alpha_1,\dots,\alpha_{n})=0$ for $i = 1, \ldots, n$, then $\alpha_1=\dots=\alpha_{n}=0$.  Because $f_n$ lies in the span of $f_1, \ldots, f_{n - 1}$ for $c \neq 0$ by Lemma \ref{lem:sum-zero}, this will imply that $\supp(A/I_c) = \{(0, \ldots, 0)\}$ and hence $A/I_c$ has dimension $0$, completing the proof.

Suppose that the $\alpha_i$ take values $\{s_1,\dots,s_{r}\}$, where $s_i$ occurs with multiplicity $m_i > 0$ so that $g(z)=\prod_i (1-s_iz)^{m_i}$, $\sum_i m_i s_i = 0$, and $\sum_i m_i = n$.  We claim that for generic $c$ this implies $s_i = 0$ for all $i$. Let $B = k[s_1, \ldots, s_r]/(\sum_i m_i s_i = 0)$; it suffices to check that with $I_c$ interpreted as an ideal in $B$, the quotient $B/I_c$ is finite dimensional over $k$.

Define the polynomial $h(z) = \prod_{i = 1}^r (z - s_i)$, and let the series expansion of $F(z)$ be given by 
\[
F(z) = \sum_{i \geq 0} a_i(s_1, \ldots, s_r, c) z^i,
\]
where $a_i$ has degree $i$ in $s_1, \ldots, s_r$ and $a_0(s_1, \ldots, s_r, c) = 1$.  Define a degree $p$ polynomial in $z$ by
\[
\wF(z; s_1, \ldots, s_r, c) := \sum_{i = 0}^p a_i(s_1, \ldots, s_r, c) z^{p - i}.
\]
In this notation, we have that 
\[
f_i = [z^p]\frac{1}{1-s_iz}F(z) = \wF(s_i; s_1, \ldots, s_r, c) = 0
\]
for $i = 1, \ldots, r$.  Notice that $\wF(s_i; s_1, \ldots, s_r, c)$ is a polynomial in $s_1, \ldots, s_r$ of homogeneous degree $p$.  By the definition of $F(z)$, we have
\[
\wF(z; s_1, \ldots, s_r, 0) = z^p,
\]
meaning that if $c = 0$, we have $s_1 = \cdots = s_r = 0$, hence $B/I_0$ is finite dimensional over $k$.

Now, let $B^d$ and $I_c^d$ denote the degree $d$ pieces of $B$ and the homogeneous ideal $I_c$, respectively.  Choose a monomial basis $\{t_i\}$ for $B^d$ independent of $c$, and let $I_c^d$ be spanned by a finite set of polynomials $\{g_j\}$ with $g_j = \sum_i h_{ji}(c) t_i$.  Notice that $\dim I_c^d$ is given by the size of the maximal non-vanishing minor of the matrix $H = (h_{ji})$.  On the other hand, we just showed that there is some degree $d > 0$ so that $I^d_0 = B^d$, meaning that $\dim I^d_c < \dim B^d$ exactly when $c$ lies in the zero set of all $\dim B^d$ minors of $H$.  This implies that for all but finitely many $c$ we have $\dim I^d_c = \dim B^d$ and hence $B/I_c$ is finite-dimensional over $k$.

For each of the finitely many choices of $r, m_1, \ldots, m_r > 0$ with $\sum_i m_i = n$, we conclude that for all but finitely many $c$, if $\{\alpha_1, \ldots, \alpha_n\}$ contains $s_i$ with multiplicity $m_i$, then $\alpha_1 = \cdots = \alpha_n = 0$.  Taking the union over excluded possibilities for $c$, we conclude that $f_i(\alpha_1, \ldots, \alpha_n) = 0$ implies $\alpha_1 = \cdots = \alpha_n = 0$ unconditionally for generic $c$, so $A/I_c$ is a complete intersection as desired.
\end{proof}

\begin{remark}
Proposition \ref{prop:ci} is a formal power series analogue of \cite[Theorem 3.2]{CE}.  However, our proof differs from the ``residues by parts'' argument which appears there, as the crucial \cite[Lemma 3.2]{CE} fails in the modular case.  It is interesting to note that our proof does not appear to translate to the characteristic $0$ case, as no analogue of the polynomial $\wF(z; s_1, \ldots, s_r, c)$ exists in that setting.
\end{remark}

@@

\begin{remark}
In the proof of Proposition \ref{prop:ci}, we require that $c$ avoids $0$ for Lemma \ref{lem:sum-zero}, $c$ avoids $\{-1, 1,\dots,(p-1)/2\}$ for Proposition \ref{prop:linind}, and that $c$ avoids a non-explicit finite set given by vanishing of a determinantal ideal.  These are the only uses of the assumption that $c$ is generic, so Proposition \ref{prop:ci} and Theorem \ref{thm:main} hold for $c$ avoiding these values.
\end{remark}

>>>>

\newcommand{\xx}[1]{\textcolor{blue}{#1}}
\newcommand{\zb}{\overline{z}}
\newcommand{\cc}[1]{\overline{#1}}
\newcommand{\m}[1]{\left|#1\right|}
\newcommand{\Res}{\operatorname{Res}}
\newcommand{\p}[1]{\left(#1\right)}
\newcommand{\rank}{\operatorname{rank}}
\newcommand{\sgn}{\operatorname{sgn}}
\newcommand{\D}{\triangle}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\f}[1]{\sum_{n=-\infty}^\infty #1_ne^{inx}}
\newcommand{\fn}[1]{\sum_{n=-\infty}^\infty #1e^{inx}}
\newcommand{\BB}{\mathcal{B}}
\newcommand{\M}{\mathcal{M}}
\newcommand{\mm}[1]{\m{\m{#1}}}
\newcommand{\ang}[1]{\left\langle #1 \right\rangle}
\newcommand{\h}{\mathfrak{h}}
\newcommand{\GL}{{\rm GL}}
\newcommand{\HH}{\mathcal{H}}
\newcommand{\arxiv}[1]{\href{http://arxiv.org/abs/#1}{{\tt arXiv:#1}}}
\newcommand{\Sym}{\operatorname{Sym}}
\renewcommand{\char}{\text{char}}
\newcommand{\sspan}{\text{span}}
\newcommand{\wF}{\widetilde{F}}
\newcommand{\supp}{\text{supp}}
\newcommand{\PP}{\mathbb{P}}

@@

The authors thank P. Etingof for suggesting the problem and for helpful discussions; the authors thank also the anonymous referee for comments leading to the streamlined proof of Proposition \ref{prop:ci} which appears here.  Some exploratory computations were done using Sage.  S.~D. was supported by the MIT Undergraduate Research Opportunities Program (UROP). Y.~S. was supported by a NSF Graduate Research Fellowship (NSF Grant \#1122374).  Both authors were also supported by NSF Grant DMS-1000113.

@@

\subsection{Proving $f_1, \ldots, f_{n-1}$ are singular vectors}

@@

Consider the homogeneous ideal $I_c = \langle f_1, \ldots, f_{n - 1} \rangle \subset A$ generated by the $f_i$.  In this section, we will show that $A/I_c$ is a complete intersection.  Recall from \cite[Example 11.8]{Har} that for a homogeneous ideal $I \subset k[X_0,\dots,X_m]$ with a minimal set of generators of size $l$, the quotient ring $k[X_0, \ldots, X_m]/I$ is a complete intersection if the closed projective subvariety of $\PP^m$ defined by $I$ has dimension $m - l$.

\begin{proposition}\label{prop:ci}
For generic $c$, the quotient $A/I_c$ is a complete intersection. 
\end{proposition}
\begin{proof}
By Proposition~\ref{prop:linind}, $I_c$ has a set of $n-1$ linearly independent and therefore minimal generators $f_1, \ldots, f_{n-1}$ in degree $p$.  Therefore $A/I_c$ is a complete intersection if and only $I_c$ cuts out a projective variety of dimension $(n - 2) - (n - 1) = -1$, meaning it is empty.  This occurs if and only if the saturation of $I_c$ is $A$, which occurs if and only if $\dim_k(A/I_c) < \infty$.

If $c = 0$, by definition of $f_i$ we see that $f_i = x_i^p$.  Note that 
\[
f_n = x_n^p = (-x_1 - \cdots - x_{n-1})^p = - x_1^p - \cdots - x_{n - 1}^p = - f_1 - \cdots - f_{n -1} \in I_0,
\]
meaning that $(x_1^p, \ldots, x_n^p) \subset I_0$ and hence $\dim_k(A/I_c) < \infty$.  Now, let $A^d$ and $I_c^d$ denote the degree $d$ pieces of $A$ and $I_c$, respectively.  Choose a monomial basis $\{t_i\}$ for $A^d$ independent of $c$, and let $I_c^d$ be spanned by a finite set of polynomials $\{g_j\}$ with $g_j = \sum_i h_{ji}(c) t_i$.  Notice that $\dim I_c^d$ is given by the size of the maximal non-vanishing minor of the matrix $H = (h_{ji})$.  On the other hand, we just showed that there is some degree $d > 0$ so that $I_0^d = A^d$, meaning that $\dim I_c^d < \dim A^d$ exactly when $c$ lies in the zero set of all size $\dim A^d$ minors of $H$.  This implies that for all but finitely many $c$ we have $\dim I_c^d = \dim A^d$, hence $\dim_k(A/I_c) < \infty$ and $A/I_c$ is a complete intersection.
\end{proof}

\begin{remark}
Proposition \ref{prop:ci} is a formal power series analogue of \cite[Theorem 3.2]{CE}.  However, our proof differs from the ``residues by parts'' argument which appears there, as the crucial \cite[Lemma 3.2]{CE} fails in the modular case.  It is interesting to note that our proof does not appear to translate to the characteristic $0$ case, as there is no analogue of operation of specializing $c$ to $0$.
\end{remark}

@@

\begin{remark}
In the proof of Proposition \ref{prop:ci}, we require that $c$ avoids $\{-1, 1,\dots,(p-1)/2\}$ for Proposition \ref{prop:linind} and that $c$ avoids a non-explicit finite set given by vanishing of a determinantal ideal.  These are the only uses of the assumption that $c$ is generic, so Proposition \ref{prop:ci} and Theorem \ref{thm:main} hold for $c$ avoiding these values.
\end{remark}
